---
title: "Tidy up data with R(Studio)"
author: "[Haiyang Jin](https://haiyangjin.github.io/)"
date: "`r format(Sys.time(), '%b %d %Y')`"
output:
  html_document:
    df_print: paged
    number_sections: true
    toc: true
    toc_depth: 6
    toc_float: 
      collapsed: true
      smooth_scroll: false
---

# Set up
```{r pre-setup}
# install tidyverse if needed
if (!require("tidyverse")) {install.packages("tidyverse")}
```

```{r setup}
# load tidyverse
# more information on tidyverse could be found: https://www.tidyverse.org
library(tidyverse)
```

# Read csv into R (What we have)
Let us read one csv file into R(Studio):
```{r message=FALSE}
# "<-" is similar to "=" in other languages
df_subj1 <- read_csv("data/238803_210821_111114 output.csv") 
head(df_subj1) # show the first six rows of df_subj1
```
The data is part of a composite word project and the independent variables (in this data file) were:  
- **Valence** (`valence`): positive vs. negative  
- **Congruency** (`congruency`): congruent vs. incongruent  
- **Alignment** (`alignment`): aligned vs. misaligned  
- (**Answer** (`corre_response`): same vs. different)  

The dependent variables were:
- **Behavioral responses** (`response`)  
- **Response times** (`RT`)

# What we want

Tidied up data could be saved as long or wide format. Typically, long format is needed in R whereas wide format is needed in SPSS, [Jamovi](https://www.jamovi.org), [JASP](https://jasp-stats.org) etc.

## Long format

Each row is one trial/participant, and columns are different variables/information.

SubjCode | Age | Gender | IV1 | IV2 | DV1 | DV2
:--: | :--: | :--: | :--: | :--: | :--: | :--:
sub-001 | 20 | M | a1 | b1 | 840 | 0.80
sub-001 | 20 | M | a2 | b1 | 800 | 0.75
sub-001 | 20 | M | a1 | b2 | 740 | 0.78
sub-001 | 20 | M | a2 | b2 | 825 | 0.85
sub-002 | 22 | F | a1 | b1 | 800 | 0.85
sub-002 | 22 | F | a2 | b1 | 700 | 0.80


## Wide format (SPSS, Jamovi...)

Each row is one participant and each column is one level (of within-participant variables). 

e.g., response times:

SubjCode | Age | Gender | a1b1 | a1b2 | a2b1 | a2b2 
:--: | :--: | :--: | :--: | :--: | :--: | :--:
sub-001 | 20 | M | 840 | 740 | 800 | 825
sub-002 | 22 | F | 800 | 650 | 700 | 875

<br>

**Question**: is `df_subj1` in a long or wide format?
```{r}
head(df_subj1, 10)
```


# What we need to do (from "what we have" to "what we want")?

0. Read data files into R;
1. Remove some rows;
2. Remove some columns;
3. Rename columns;
4. Reorder/re-arrange columns;
5. Remove trials whose RT is too short (or "too long");
6. Calculate means of each condition for each participant;
7. Remove participants who responded "too slow" or "too fast";
8. Conversion between long and wide formats;
9. Save the final data as csv files;
10. ...

# Load data

```{r}
# filenames for all csv files in data/
allcsv <- list.files(path="data/", pattern="*.csv")
head(allcsv)
```

```{r message=FALSE}
# read all files at once
# read_csv("data/551155_210821_114658 output.csv")  # what we used earlier for reading a single file
df_all <- map_dfr(allcsv, ~read_csv(file.path("data", .x)))
head(df_all)
```

```{r}
# filter out trials with "space" response, which were not our interests
df_a_filtered <- filter(df_all, response != "space")
head(df_a_filtered)
```


```{r}
# create a subjcode column
df_af_subj <- mutate(df_a_filtered,
                     subjcode = as.integer(as_factor(prolific_num)))
head(df_af_subj)
```

# Pipe (%>%)

```{r message=FALSE}
# comparable codes to the above steps
df_tmp <- list.files(path="data/", pattern="*.csv") %>% 
  # map_dfr(., ~read_csv(file.path("data", .x))) %>% 
  map_dfr(~read_csv(file.path("data", .x))) %>% 
  filter(response != "space")  %>% 
  mutate(subjcode = as.integer(as_factor(prolific_num))) 
head(df_tmp)
```

```{r message=FALSE}
df_raw <- list.files(path="data/", pattern="*.csv") %>% # all data file names
  map_dfr(~read_csv(file.path("data", .x))) %>% # read all these files
  filter(response != "space")  %>%  # remove trials whose responses were "space"
  mutate(subjcode = as.integer(as_factor(prolific_num)), # make participant code
         answer = corre_response) %>% # rename corre_response as answer
  # only select columns we want
  select(subjcode, valence, congruency, alignment, answer, response, RT, correct)

head(df_raw, 10)
```

# Remove outliers

- At the trial level
  - Only keep trials whose RT were longer than 200 ms;
  - Only keep trials whose RT were within three SD (for each participant separately);
- At the participant level
  - Only keep participants where the ratio of removed trials were below 15%;
  - Only keep participants whose Accuracy is above 65%;
  - Only keep participants whose RT is within three SD (across participants);

## At the trial level
Trials whose RT were not longer than 200 ms or outside **three** standard deviations within each participant were removed.
```{r }
df_3sd <- df_raw %>% 
  filter(RT > 200) %>% # only keep trials whose RT is longer than 200
  group_by(subjcode) %>% # perform following calculation for each participant separately
  mutate(z_RT = scale(RT),  # calculate Z-value for each participant separately
         N_total = n()) %>% # number of trials in each group
  filter(z_RT >= -3 & z_RT <= 3) %>% 
  mutate(N_remain = n()) %>%  # number of trials within 3 SD 
  ungroup() # remove any grouping
```


## At the participant level

### Ratio of removed trials
```{r }
df_remove <- df_3sd %>% 
  select(subjcode, N_total, N_remain) %>% 
  distinct() %>% # find all the unique rows
  mutate(N_remove = N_total - N_remain,
         N_ratio = N_remove/N_total)

df_remove %>% 
  filter(N_ratio > 0.15)
```
For all participants, the ratio of outlier trials were below 15%. Therefore, no participants were removed here.

```{r}
df_remove %>% 
  summarize(mean_N_remove = mean(N_remove),
            mean_N_ratio = mean(N_ratio))
```
In summary, 5.92 trials were removed for each subject on average, which were about 1.5% of all trials.

### Overall accuracy
Participants whose overall accuracy was below 65% were removed.
```{r}
df_acc <- df_3sd %>%
  group_by(subjcode) %>% # perform following calculation for each participant separately
  summarize(mean_acc = mean(correct)) # mean accuracy

df_subj_65 <- filter(df_acc, mean_acc >= .65)

df_acc %>% 
  filter(mean_acc < 0.65) 
```
As a result, one participant was removed due to its overall accuracy below 65%.

### Response times within three SD
Participants whose mean response times (RT) were outside **three** standard deviations among participants were removed.
```{r }
df_subj <- df_3sd %>% 
  group_by(subjcode) %>% # perform following calculation for each participant separately
  summarize(meanRT = mean(RT)) %>% # calculate the mean RT 
  mutate(z_meanRT = scale(meanRT))  # calculate Z-value for each participant separately

df_subj_3SD <- filter(df_subj, z_meanRT >= -3 & z_meanRT < 3 )

df_subj %>% 
  filter(z_meanRT > 3 | z_meanRT < -3)  # show subjects outside 3SD
```
One participant was removed due to mean RT being outside 3SD.

## Number of remaining participants
```{r}
# participant list after removing outlier (meet criteria for both accuracy and RT)
subj_list <- intersect(unique(df_subj_65$subjcode),
                       unique(df_subj_3SD$subjcode))

# only keep subjects who are in subj_list
df_clean <- df_3sd %>% 
  filter(subjcode %in% subj_list)
head(df_clean, 10)
```

Number of participants for each experiment (after removing participants): 
```{r}
df_clean$subjcode %>% # all values in subjcode
  unique() %>% # find all unique values
  length() # how many unique values
```
In summary:  
One participant was removed in total (due to mean RT being outside 3SD and accuracy being below 65%).

# "Clean" data
## Long format

Each row is one trial:
```{r}
df_clean
```

Each row is one participant and each column is one variable (could be independent or dependent variables):
```{r}
# Accuracy for each condition
df_long_acc <- df_clean %>% 
  # perform following steps in each group separately
  group_by(subjcode, valence, congruency, alignment) %>% 
  summarize(acc = mean(correct), # accuracy
            N_trial = n(),       # number of trials in each condition
            .groups = "drop")
head(df_long_acc)
```

```{r}
# (Correct) response times for each condition
df_long_rt <- df_clean %>% 
  filter(correct == 1) %>% # only keep correct trials
  group_by(subjcode, valence, congruency, alignment) %>% 
  summarize(meanRT = mean(RT), # mean correct response times
            N_trial = n(),     # number of trials in each group
            .groups = "drop")
head(df_long_rt)
```

## Wide format
Each row is one participant and each column is one level of the combined independent variables:
```{r}
# convert Accuracy from long into wide format
df_wide_acc <- df_long_acc %>%
  pivot_wider(id_cols = subjcode,
              names_from = c(valence, congruency, alignment), # independent variables
              names_sep = "_",
              values_from = acc) # dependent variables
df_wide_acc
```

```{r}
# convert correct response times from long into wide format
df_wide_rt <- df_long_rt %>% 
  pivot_wider(id_cols = subjcode,
              names_from = c(valence, congruency, alignment), # independent variables
              names_sep = "_",
              values_from = meanRT) # dependent variables
df_wide_rt
```

# Save as csv file
```{r}
# save the wide data as csv
write_csv(df_wide_acc, "acc_wide.csv")
write_csv(df_wide_rt, "rt_wide.csv")
```

# Summary

0. Read data files into R (`read_csv()`; `map_dfr()`);
1. Remove some rows (`filter()`);
2. Remove some columns (`select()`);
3. Rename columns (`mutate()`);
4. Reorder/re-arrange columns (`select()`);
5. Remove trials whose RT is too short (or "too long") (`scale()`; `filter()`);
6. Calculate means of each condition for each participant (`group_by()`; `summarize()`);
7. Remove participants who responded "too slow" or "too fast" (`scale()`; `filter()`);
8. Conversion between long and wide formats (`pivot_longer()`; `pivot_wider()`);
9. Save the final data (`write_csv()`);
10. Help documents (e.g., `?write_csv`) and [tidyverse website](https://www.tidyverse.org).

# Versions of packages used
```{r versions}
sessionInfo()
```
